Supporting Information: Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes.

### 2020-05-14 Ben_Manuscripts/hdphmm/Supporting_Information.tex
New supporting figures and mods to draft

  \subsection{Tips for Reliable State Sequence Identification}\label{section:fitting_tips}

@@

>>>>

  \section{Tips for Reliable State Sequence Identification}\label{section:fitting_tips}

@@

  \section{Deviations of the IHMM from Molecular Motion}\label{section:shorttimes_msd}

   At extremely short time lags, the MSD is generally over-predicted by the IHMM
  (see Figure~\ref{fig:unclustered_msd_shortlag}) because the VAR(1) model assumes
  multivariate Gaussian noise while the MD data suggests the noise is better 
  modeled by more general L\'evy stable noise. The first time lag of the MSD is
  equivalent to the variance of all observed fluctuations (first-order difference). 
  The IHMM simulation appears to over-estimate the variance because the actual 
  distribution of solute fluctuations has heavy tails with a smaller variance (Figure~\ref{fig:emission_widths}).
  The total distribution of hops is expected to be heavy tailed because it is a
  combination of hop distributions from many states with different variances. More
  interestingly, in Figure ~\ref{fig:state_emission_widths}, we show that the 
  emission distribution of a single state picked out by the IHMM is fit better by
  a more general L\'evy distribution while the emissions from IHMM realizations are
  Gaussian, as expected. An improvement to the IHMM that may allow dynamics more 
  faithful to MD would use an autoregressive model with L\'evy stable noise. However,
  this would be less computationally efficient due to lack of conjugate priors.

  \begin{figure}
  \centering
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{unclustered_msd_MET_shortlag.pdf}
  \caption{}\label{fig:unclustered_msd_shortlag}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{emission_widths.pdf}
  \caption{}\label{fig:emission_widths}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
  \includegraphics[width=\textwidth]{state_emission_widths.pdf}
  \caption{}\label{fig:state_emission_widths}
  \end{subfigure}
  \caption{(a) At short time lags, the MD MSD (black) and the IHMM MSD (orange)
  differ significantly in their shape and curvature. (b) While the 
  fluctuations in $z$ of both distributions are heavy-tailed, fitting well to zero-centered
  symmetric L\'evy stable distributions with stability parameters, $\alpha$, the IHMM has
  a wider variance. (c) The distribution of fluctuations in a chosen state frequently
  visited by methanol is heavy tailed, but the IHMM is constrained to produce fluctuations
  from a multivariate normal distribution ($\alpha$=2 in each dimension).
  }\label{fig:short_timelags}
  \end{figure}

  Also apparent from Figure~\ref{fig:unclustered_msd_shortlag} is that the 
  curvature of the MD MSD is far more gradual than that generated by the IHMM.
  This is because the VAR(1) model is only correlated to its previous fluctuation
  while the MD simulations suggest that correlation persists for at least tens
  of nanoseconds.
    % BJC2: could probably use a correlation function to suggest the AR order and stick the plot in SI
  One may be able to reproduce the curvature at short time lags with higher order
  VAR models, but it becomes a much higher dimensional parameterization that would
  require significantly more data in order to produce stable results. Since we 
  are interested in reproducing longer time scale behavior, it is not an issue of
  great concern.

  \section{Influence of number of clusters on qualitative hybrid trajectories}

  \begin{figure}
  \begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\textwidth]{qualitative_clustered_MET_20.pdf}
  \caption{20 clusters}\label{fig:qualitative_20}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\textwidth]{qualitative_clustered_MET_30.pdf}
  \caption{30 clusters}\label{fig:qualitative_30}
  \end{subfigure}
  \caption{The quality of hybrid trajectories resulting from clustered parameters
  sets qualitatively improves as the number of clusters is increased. This comes 
  at the cost of a larger state space to interpret.}\label{fig:qualitative_improvement}
  \end{figure}
